---
title: "Sales Forecast, Example"
subtitle: "Practice exercise: Seasonal Naive Method, Exponential Smoothing Model and ARIMA Model"
author: "JDM"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Sales Forecast Considerations:

- This file shows the result of "Sales_per_day_Forecast.R" file. It is the same code.
- Although RMarkdown idea is to hide the script, some important scripts will be shown.


This EXAMPLE shows in a simple way how to use some popular methods for Forecast:
Benchmark Method (Seasonal Naive Method), Exponential Smoothing Model and ARIMA Model.

This data is not adjusted for inflation. So these are nomimal values. (practice excersice)
Monthly retail sales per day

```{r pressure, echo=FALSE, warning=FALSE, message=FALSE}
library(fpp2)
library(ggplot2)
library(readxl)
```



```{r, echo=TRUE, warning=FALSE, message=FALSE}
# Load data
data <- read.csv("https://raw.githubusercontent.com/Jdmuci/R_practices/main/sales_per_day.csv")

# declare as time series data
Y <- ts(data[,2], start = c(1992,1), end = c(2017,12), frequency = 12)
```


## Preliminary analysis

Time plot

```{r, echo=FALSE, warning=FALSE, message=FALSE}
autoplot(Y) +
  ggtitle("Time plot: US Retail Sales Per Day") +
  ylab("Millions of dollars")
```

### 1- Seasonality and Trend

Data has a strong trend and seasonality. Need to investigate transformations.

- Take the first difference to remove trend.

```{r, echo=TRUE, warning=FALSE, message=FALSE}
DY <- diff(Y)
```

Time plot of differenced data:

```{r, echo=FALSE, warning=FALSE, message=FALSE}
autoplot(DY) +
  ggtitle("Time plot: Change in US Retail Sales Per Day (1st difference)") +
  ylab("Millions of dollars")
```

Series appears trend-stationary, but need to investigate seasonality.

Lets now focus on seasonality. The next two plots are interesting and useful to analyse seasonality.


```{r, echo=FALSE, warning=FALSE, message=FALSE}
ggseasonplot(DY) +
  ggtitle("Seasonal plot: Change in Daily Retail Sales") +
  ylab("Millions of dollars")
# This function makes a series for each year and plot them together in a 12 months period to see the seasonality.
```

This function makes a series for each year and plot them together in a 12 months period to see the seasonality. 

Apparently there is a drop between december and january, and a rise from november to december (christmas season).

Lets look at another seasonal plot, the subseries plot:

```{r, echo=FALSE, warning=FALSE, message=FALSE}
ggsubseriesplot(DY)
# This function graphs each month of the series separately. Example: every January together. Average: blue line.
```

This function graphs each month of the series separately. Example: every January together (Ejemplo: todos los meses Enero de la serie, juntos) . Average: blue line.


So, we confirm the fluctuations we saw in the first plot is seasonality every december and january.


## Forecast with various methods (considering seasonality).

## FIRST METHOD:    Use a Benchmark method to forecast. (there are many)

So, lets use the Seasonal naive method as our benchmark:
y_t = y_{t-s} + e_t

Using the differenced series (1st difference):

```{r, echo=FALSE, warning=FALSE, message=FALSE}
fit <- snaive(DY)  # Residual SD = 231.45, this is our benchmark
print(summary(fit))
checkresiduals(fit)
```

We see residuals ACF are not ideal, lets look if we can find a better model

### SECOND METHOD:    Fit ETS method
Another type we can try is Exponential smoothing model.

They (exponential smoothing model) use weighted averages of past observations to forecast new values
Exponential smoothing models are based on a description of the trend and seasonality in the data

We can use regular data (not the differenced one)

```{r, echo=FALSE, warning=FALSE, message=FALSE}
fit_ets <- ets(Y)          # Residual SD = 0.0197
print(summary(fit_ets))
checkresiduals(fit_ets)
```

Apparently it did not improve, still correlation left in the data and residuals look almost the same
But it fits better, since the  Residual SD is much smaller.

 Lets do another type of model:
 
### THIRD METHOD:   Fit an ARIMA Model

Exponential smoothing and ARIMA models are the two most widely used approaches to time series forecasting.

- ARIMA models aim to describe the autocorrelation in the data.
- Needs to be stationary data

So lets use differenced data, there is seasonality.


 We can use regular data and tell it to take 1st difference (d=1) and 1st seasonality difference (D=1).
 
- "stepwise = FALSE" finds the model that fits better and try every single possibility, so it could take a lot of time
- "aproxximation = FALSE" will show the exactly result, not an approximation
- "Trace" will print all the models 

```{r, echo=FALSE, warning=FALSE, message=FALSE}
fit_arima <- auto.arima(Y, d=1, D=1, stepwise = FALSE, approximation = FALSE, trace = FALSE)  # Residual SD = 174.89
print(summary(fit_arima))
checkresiduals(fit_arima)
```

```{r, echo=TRUE, warning=FALSE, message=FALSE}
#SD = sqrt(sigma^2)
sqrt(30589)
```

This model fits better, there is not much autorrelation in residuals (ACF), but is not perfect!
 
It means there is better model for this time series, but lets use this model.


##     FORECAST 


### Forecast with ARIMA Model

```{r, echo=TRUE, warning=FALSE, message=FALSE}
fcst <- forecast(fit_arima, h=24)
autoplot(fcst, include = 60)
print(summary(fcst))
```


It looks realistic 
 - "h = 24" forecasted period (months)
-  function Include, means the number of months (historic data)
